Scaling of Electrode-Electrolyte Interface Model Parameters In Phosphate Buffered Saline

### 2014-04-03 paper.tex
Added preliminary responses to reviewer comments

Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, i.e. a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution.\cite{Merrill05} This capacitive behaviour may also be a result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}
This displacement mechanism is capable of transferring charge between electrical and biological systems without damaging electrodes or tissue.\cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE); also known as a fractional capacitor.
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} where the value of each R-C pair is chosen to set the slope of impedance relative to frequency. The placement of R-C pairs within the CPE is determined by Morrison's parameters $m$ and $k$, where $k$ determines the spacing, or density, of R-C branches and $m$ determines the resulting slope.
We measure displacement currents by sweeping the frequency of a sinusoidal stimulus across electrodes 2 and 7 of the Octrode while measuring the voltage between electrodes 2 and 3 (numbering shown in Fig.~\ref{fig:octrode}).
This allows for the measurement of a single interface impedance in series with the resistance presented by the bulk solution.
\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionVsFrequency_magnitude}
    \end{center}
    \caption{Charge injected per phase during CPE measurement for each of the six concentrations of PBS. Calculated using a geometric surface area of 10\thinspace mm$^{2}$}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}

@@

Our model uses two reverse connected diodes to represent Faradaic current conduction between electrode and electrolyte. Specifically we use the diodes to model irreversible Faradaic reactions. (Any reversible reactions, those producing reactants bound to the surface of the electrode, are encapsulated by the displacement branch (CPE) of the model.)
Possible irreversible Faradaic reactions for platinum electrodes in saline are:
\begin{eqnarray}
    Pt + 4Cl^{-} &\Rightarrow& [PtCl_{4}]^{2-} + 2 e^{-}\\
    2H_{2}O + 2 e^{-} &\Rightarrow& H_{2}\uparrow + 2OH^{-}\\
    2H_{2}O &\Rightarrow& O_{2}\uparrow + 4H^{+} + 4e^{-}\\
    2Cl^{-} &\Rightarrow& Cl_{2}\uparrow + 2e^{-}\\
    Cl^{-} + H_{2}O &\Rightarrow& ClO^{-} + 2H^{+} + 2e^{-}
\end{eqnarray}

@@

@@

@@

@@

>>>>

Displacement currents are responsible for capacitive behaviour at interfacial boundaries.
They are brought about by the redistribution of charge in response to applied fields, i.e. a charged electrode will attract or repel $Cl^{-}$ from its surface and reorientate polar molecules in the surrounding solution.\cite{Merrill05} This capacitive behaviour may also be a result of electrode polarisation in the form of reversible Faradaic reactions at the surface of the electrode.
Such reactions involving water and platinum, as identified in \cite{Horch2004,Mohtashami2011,Merrill05}, include:
\begin{eqnarray}
    Pt + H_{2}O &\Leftrightarrow& PtO + 2 H^{+} + 2 e^{-}\\
    PtO + H_{2}O &\Leftrightarrow& PtO_{2} + 2 H^{+} + 2e^{-}\\
    Pt + H^{+} & \Leftrightarrow & Pt-H\\
    Pt + H_{2}O + e^{-} &\Leftrightarrow& Pt-H+OH^{-}
\end{eqnarray}

This displacement mechanism is capable of transferring charge between electrical and biological systems without damaging electrodes or tissue.\cite{Horch2004}
This behaviour is modelled by way of a Constant Phase Element (CPE); also known as a fractional capacitor.
For simulation purposes this element is realised as an array of R-C branches.~\cite{ScottSingle2013,Morrison59,Elwakil10} where the value of each R-C pair is chosen to set the slope of impedance relative to frequency. The placement of R-C pairs within the CPE is determined by Morrison's parameters $m$ and $k$, where $k$ determines the spacing, or density, of R-C branches and $m$ determines the resulting slope.
We measure displacement currents by sweeping the frequency of a sinusoidal stimulus across electrodes 2 and 7 of the Octrode while measuring the voltage between electrodes 2 and 3 (numbering shown in Fig.~\ref{fig:octrode}).
This allows for the measurement of a single interface impedance in series with the resistance presented by the bulk solution.
\begin{figure}
    \begin{center}
        \includegraphics{graphics/chargeInjectionVsFrequency_magnitude}
    \end{center}
    \caption{Charge injected per phase during CPE measurement for each of the six concentrations of PBS. Calculated using a geometric surface area of 10\thinspace mm$^{2}$}
    \label{fig:chargeInjectionVsFrequency}
\end{figure}

@@

Our model uses two reverse connected diodes to represent Faradaic current conduction between electrode and electrolyte. Specifically we use the diodes to model irreversible Faradaic reactions. (Any reversible reactions, those producing reactants bound to the surface of the electrode, are encapsulated by the displacement branch (CPE) of the model.)
Possible irreversible Faradaic reactions for platinum electrodes in saline are:

\begin{align}
    Pt + 4Cl^{-} &\Rightarrow& [PtCl_{4}]^{2-} + 2 e^{-} \label{eqn:ptCl}\\
    2H_{2}O + 2 e^{-} &\Rightarrow& H_{2}\uparrow + 2OH^{-} \label{eqn:H20}\\
    2H_{2}O &\Rightarrow& O_{2}\uparrow + 4H^{+} + 4e^{-} \label{eqn:2H20}\\
    2Cl^{-} &\Rightarrow& Cl_{2}\uparrow + 2e^{-} \label{eqn:Cl} \\
    Cl^{-} + H_{2}O &\Rightarrow& ClO^{-} + 2H^{+} + 2e^{-} \label{eqn:ClH20}
\end{align}

{
    \color{blue}
    It should be noted that each of these reactions (\ref{eqn:ptCl}, \ref{eqn:H20}, \ref{eqn:2H20}, \ref{eqn:Cl} and \ref{eqn:ClH20}) have different reaction potentials. It is expected that the situation in-vivo will be much more complex.
}

@@

    {
        \color{Blue}
        The diodes are there to model the change in current attributed to the first biologically damaging reaction that occurs at the electrode/electrolyte interface. Whatever reaction occurs first in the relevant electrolyte is the reaction that diode aims to mimic. We observe that the current increases near exponentially around this point so we use a diode to mimic that. We expect this reaction may not be the same in-vivo as it is in-vivo but the properties of a diode can be altered to match where this occurs.

        If there are secondary reactions occurring at higher potentials then we could add a second pair of diodes with the properties of that specific reaction. The problem here is that we would then need to include memristors (as used in ScottSingle2013) to account for reactant depletion. We don't account for reactant depletion for the same reason we don't account for reactions other than the very first Faradaic reaction to occur. The reason is that we never want biologically damaging reactions to occur due to an implant, so if they happen then something has gone wrong.

        Think of our attempt to model Faradaic reactions as a ``you've pushed this too far'' indicator for an electrical engineer.

        I think you're essentially right with you're argument, we're not up-front about this and we probably portray more capability in the Faradaic modelling aspect than we are able or aim to achieve.
    }

@@

    {
        \color{Blue}
        Correct, we do not wish to determine exactly what is happening at the electrode surface - we want to create a flexible model that can be used to simulate impedance.

    }

@@

    {
        \color{Blue}
        This is a good point. ScottSingle2013 compare the model prediction to stimulus waveforms. Our aim is to see what changes in that model as the concentration of PBS is varied and to compare those measurements to in-vivo measurements using a live sheep. Were doing this because 0.1X PBS is commonly used as a phantom for real spine. There hasn't been an objective way of determining how accurate that assumption was until we used this model to make that comparison.
    }

@@

    {
        \color{Blue}
        We need this because it represents the resistance of the interface itself, separate from that of the fluid.
        We measure the impedance magnitude of the displacement currents at high frequency (flattened part of Fig. 5). For each concentration the impedance magnitude at 10kHz is higher than is predicted by the transimpedance measurements made in Fig. 3. The transimpedance measurments do not measure impedance across an interface that is passing current and therefore any voltage drop across an interface due to $R_{S}$ would vanish. The displacement measurements take the voltage across interfaces that are passing stimulus currents. These measurements see the potential drop across the interface due to the current flow via Ohms law.

        The transimpedance measurements give us an interelectrode resistance. The displacement measurements at high frequency give us an interelectrode impedance plus an interface impedance. Subtracting one from the other gives us a value of resistance that only appears as we pass current through an interface, which we include as $R_{S}$.
    }
